### 2017-11-15 Draft2/Appending.Compare.tex
Many edits

In this section, $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ will denote the write cost functions for dynamic and growth arrays, respectively. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
From Theorem \ref{thm:BothSidesInequality}, I may divide all sides of this inequality by $n$ to receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
It was shown earlier that $\FuncWritesGrowthArray(n) \sim n$. From Theorem \ref{thm:MergesOverOps}, $\dfrac {\FuncWritesDynamicArray(n)} n \sim \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. From Theorem \ref{thm:InterchangeableInInequality}, it follows that
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
\HdrSpaceComplexCmp

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space cost functions for dynamic and growth arrays, respectively. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and that $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Taking the ratio of $\FuncSpaceDynamicArray(n)$ to $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \frac{1}{\VarGrowthFactor^\VarUseful\VarInitCapacity} \cdot \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity &&\text{(Theorem \ref{thm:MergesOverOps})}\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:RemovesLowerOrderTerms})}
\end{align*}

>>>>

In this section, $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ will denote the write cost functions for dynamic and growth arrays, respectively. My goal is to approximate the ratio $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$ for large $n$. From my assumption near the beginning of Section \ref{subsec:Appending}, this is equal to $\dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
From Theorem \ref{thm:BothSidesInequality}, I may divide all sides of this inequality by $n$ to receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
It was shown earlier that $\FuncWritesGrowthArray(n) \sim n$. From Theorem \ref{thm:MergesOverOps}, $\dfrac {\FuncWritesDynamicArray(n)} n \sim \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. From Theorem \ref{thm:InterchangeableInInequality}, it follows that
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``For large $n$, dynamic arrays make approximately $2$ to $3$ times as many writes as growth arrays. \underline{Thus, for large $n$, they are about $2$ to $3$ times slower than growth arrays.}''

\HdrSpaceComplexCmp

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space cost functions for dynamic and growth arrays, respectively. I wish to approximate the ratio $\dfrac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)}$ for large $n$. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and that $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Dividing $\FuncSpaceDynamicArray(n)$ by $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \frac{1}{\VarGrowthFactor^\VarUseful\VarInitCapacity} \cdot \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity &&\text{(Theorem \ref{thm:MergesOverOps})}\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:RemovesLowerOrderTerms})}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``\underline{For large $n$, dynamic arrays use approximately twice as much space as growth arrays.}''

### 2017-11-15 Draft2/Appending.Dynamic.tex
Many edits

I will na\"{i}vely assume that $\FuncTime(n) \propto \FuncWrites(n)$, or that the average time for $n$ appends is proportional to their write cost. Then $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)} = \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$, where $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ are respectively the write cost functions for dynamic and growth arrays. In this section, I will derive the formula for $\FuncWritesDynamicArray(n)$. Because this section concerns dynamic arrays only, I will denote it $\FuncWrites(n)$.

>>>>

I will assume that $\FuncTime(n) \propto \FuncWrites(n)$, or that the average time for $n$ appends is proportional to their write cost (although this is an oversimplification). Then $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)} = \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$, where $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ are respectively the write cost functions for dynamic and growth arrays. In this section, I will derive the formula for $\FuncWritesDynamicArray(n)$. Because this section concerns dynamic arrays only, I will denote it $\FuncWrites(n)$.

### 2017-11-15 Draft2/Appending.Growth.tex
Many edits

% Bookmark
% The following following paragraph doesn't work.

Now, I wish to approximate this using $\sim$. Since $\VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$, it follows from Lemma \ref{lem:ToVarUsefulPowerBounds} that $n \FluteLeq \VarCapacitySeq_\VarUseful \FluteLess \VarGrowthFactor{n}$. Since $n$ and $\VarGrowthFactor{n}$ are both $O(n)$, [the squeeze theorem] implies that $\VarCapacitySeq_\VarUseful = O(n)$.

Now, consider the ratio $\ExprNToInfty \dfrac {\FuncSpaceTail(\VarUseful)} {\VarCapacitySeq_\VarUseful} = \dfrac {O(\log n)} {O(n)} = O \left( \dfrac{\log n}{n} \right) = 0$. Because it is $0$, from Theorem \ref{thm:RemovesLowerOrderTerms} one can conclude
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful) \sim \VarCapacitySeq_\VarUseful
\end{align*}
Substituting $\FuncSpace(n)$ for $\VarCapacitySeq_\VarUseful$ in the previous inequality,
\begin{align*}
n \FluteLeq \FuncSpace(n) \FluteLess \VarGrowthFactor{n}
\end{align*}

>>>>

Now, I wish to approximate this using $\sim$. Since $\VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$, it follows from Lemma \ref{lem:ToVarUsefulPowerBounds} that $n \FluteLeq \VarCapacitySeq_\VarUseful \FluteLess \VarGrowthFactor{n}$. From Corollary \ref{coro:BothSidesInequalityInvert}, $\dfrac{1}{\VarGrowthFactor n} \FluteLess \dfrac{1}{\VarCapacitySeq_\VarUseful} \FluteLeq \dfrac{1}{n}$. From Theorem \ref{thm:BothSidesInequality}, $\dfrac{\FuncSpaceTail(\VarUseful)}{\VarGrowthFactor n} \FluteLess \dfrac{\FuncSpaceTail(\VarUseful)}{\VarCapacitySeq_\VarUseful} \FluteLeq \dfrac{\FuncSpaceTail(\VarUseful)}{n}$.

Since $\FuncSpaceTail(\VarUseful) = O(\VarUseful) = O(\log n)$, the limits of the lower and upper bounds are both $0$. From the squeeze theorem, it follows that $\ExprNToInfty \dfrac{\FuncSpaceTail(\VarUseful)}{\VarCapacitySeq_\VarUseful} = 0$. Thus, $\FuncSpaceTail(\VarUseful) = o(\VarCapacitySeq_\VarUseful)$. From Theorem \ref{thm:RemovesLowerOrderTerms}, one can conclude
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful) \sim \VarCapacitySeq_\VarUseful
\end{align*}
Using Theorem \ref{thm:InterchangeableInInequality}, I may substitute $\FuncSpace(n)$ for $\VarCapacitySeq_\VarUseful$ in the first inequality to receive
\begin{align*}
n \FluteLeq \FuncSpace(n) \FluteLess \VarGrowthFactor{n}
\end{align*}
This result is significant since it has smaller lower and upper bounds than $\FuncSpaceDynamicArray(n)$. Since the same values of $n$ as for $\FuncSpaceDynamicArray(n)$ correspond to when it is close to the bounds for $\FuncSpace(n)$ (namely, when $n$ is approximately $\VarGrowSeq_i$), $\FuncSpace(n)$ is asymptotically less than $\FuncSpaceDynamicArray(n)$ even though their intervals overlap. This will be proven rigorously in the space complexity comparison section.

### 2017-11-15 Draft2/Appendix.tex
Many edits

{\HdrNote} In the following theorems and their proofs, variables with the letters $f$, $g$, or $h$ denote functions that, for sufficiently large $n$, are positive and non-decreasing.

@@

>>>>

{\HdrNote} In the following theorems and their proofs, variables with the letters $f$, $g$, or $h$ denote functions that are positive for sufficiently large $n$.

@@

\subsubsection{$\FluteLess$ and $\FluteLeq$ Inequalities are Flipped when Both Sides are Inverted}

\begin{repcorollary}{coro:BothSidesInequalityInvert}
	If $f \FluteLess g$, then $\dfrac{1}{g} \FluteLess \dfrac{1}{f}$.\\
	This statement also holds for $\FluteLeq$.
\end{repcorollary}

\begin{proof}
	This immediately follows from $f \FluteLess g$ or $f \FluteLeq g$ by using Theorem \ref{thm:BothSidesInequality} to divide both sides by $fg$.
\end{proof}

### 2017-11-15 Draft2/AsymptoticNotation.tex
Many edits

{\HdrNote} In the following theorems, variables with the letters $f$, $g$, or $h$ denote functions that, for sufficiently large $n$, are positive and non-decreasing.

@@

>>>>

{\HdrNote} In the following theorems, variables with the letters $f$, $g$, or $h$ denote functions that are positive for sufficiently large $n$.

@@

\begin{corollary}
\label{coro:BothSidesInequalityInvert}
	If $f \FluteLess g$, then $\dfrac{1}{g} \FluteLess \dfrac{1}{f}$.\\
	This statement also holds for $\FluteLeq$.
\end{corollary}

### 2017-11-15 Draft2/FieldsProperties.tex
Work
Make algorithms not float
Add pictures, [WIP] title all the algos
Many edits

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns the capacity of $\VarList$}
		\State $\VarList.\FieldCapacity := \VarList.\FieldBuffer.\FieldLength$
		\State
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\Procedure{$\FuncConstructor$}{$\VarList$}
			\State $\VarList.\FieldBuffer \gets \FuncNewArray(\VarInitCapacity)$
			\State $\VarList.\FieldSize \gets 0$
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\begin{itemize}
	\item $\VarList.\FieldHead$ -- The \textbf{head} of $\VarList$. It returns the buffer new items are appended to.
	\item $\VarList.\FieldTail$ -- The \textbf{tail} of $\VarList$. (\textbf{Important note: The tail is a dynamic array.}) It returns a dynamic array of pointers to buffers that are already filled with items. The tail can be thought of as a two-dimensional array.\\
	({\HdrRemark} It may seem strange for a growth array to use the very data structure it is replacing. Lemma \ref{lem:VarUsefulIsOLogN} will show, however, that only $O(\log n)$ many pointers are appended to the tail. Thus, the extra copying and allocations the tail performs is minuscule compared to other work done by the growth array.)
	\item $\VarList.\FieldSize$ -- The number of items in $\VarList$.
	\item $\VarList.\FieldCapacity$ -- The \textbf{capacity} of $\VarList$. It returns the maximum number of items $\VarList$ can hold before it must grow.
\end{itemize}

@@

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
		\State
		\LineComment{Returns the capacity of $\FieldHead$}
		\State $\VarList.\FieldHeadCapacity := \VarList.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the size of $\FieldHead$}
		\LineComment{\textbf{Explanation:} $\FieldCapacity - \FieldHeadCapacity$ is the number of items in $\FieldTail$.}
		\LineComment{$\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items not in $\FieldTail$ (meaning, in $\FieldHead$).}
		\State $\VarList.\FieldHeadSize := \VarList.\FieldSize - \left(\VarList.\FieldCapacity - \VarList.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\Procedure{$\FuncConstructor$}{$\VarList$}
			\State $\VarList.\FieldHead \gets \FuncNewArray(\VarInitCapacity)$
			\State $\VarList.\FieldTail \gets \FuncNewDynamicArray()$
			\State $\VarList.\FieldSize \gets 0$
			\State $\VarList.\FieldCapacity \gets \VarInitCapacity$
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}[H]
	\caption{Properties \TextDynamicArray}
	\begin{algorithmic}
		\LineComment{Returns the capacity of $\VarList$}
		\State $\VarList.\FieldCapacity := \VarList.\FieldBuffer.\FieldLength$
		\State
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[H]
	\caption{Constructor \TextDynamicArray}
	\begin{algorithmic}
		\Procedure{$\FuncConstructor$}{$\VarList$}
			\State $\VarList.\FieldBuffer \gets \FuncNewArray(\VarInitCapacity)$
			\State $\VarList.\FieldSize \gets 0$
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\begin{itemize}
	\item $\VarList.\FieldHead$ -- The \textbf{head} of $\VarList$. It returns the buffer new items are appended to.
	\item $\VarList.\FieldTail$ -- The \textbf{tail} of $\VarList$. (\underline{{\HdrNote} The tail is a dynamic array.}) It returns a dynamic array of pointers to buffers that are already filled with items. The tail can be thought of as a two-dimensional array.\\
	({\HdrRemark} It may seem strange for a growth array to use the very data structure it is replacing. Lemma \ref{lem:VarUsefulIsOLogN} will show, however, that only $O(\log n)$ many pointers are appended to the tail. Thus, the extra copying and allocations the tail performs is minuscule compared to other work done by the growth array.)
	\item $\VarList.\FieldSize$ -- The number of items in $\VarList$.
	\item $\VarList.\FieldCapacity$ -- The \textbf{capacity} of $\VarList$. It returns the maximum number of items $\VarList$ can hold before it must grow.
\end{itemize}

@@

\begin{algorithm}[H]
	\caption{Properties \TextGrowthArray}
	\begin{algorithmic}
		\LineComment{Returns whether $\VarList$ is full}
		\State $\VarList.\FieldFull := \VarList.\FieldSize = \VarList.\FieldCapacity$
		\State
		\LineComment{Returns the \textbf{head capacity}, or the capacity of $\FieldHead$}
		\State $\VarList.\FieldHeadCapacity := \VarList.\FieldHead.\FieldLength$
		\State
		\LineComment{Returns the \textbf{head size}, or the size of $\FieldHead$}
		\LineComment{\textbf{Explanation:} $\FieldCapacity - \FieldHeadCapacity$ is the number of items in $\FieldTail$.}
		\LineComment{$\FieldSize - \left( \FieldCapacity - \FieldHeadCapacity \right)$ is the number of items not in $\FieldTail$ (meaning, in $\FieldHead$).}
		\State $\VarList.\FieldHeadSize := \VarList.\FieldSize - \left(\VarList.\FieldCapacity - \VarList.\FieldHeadCapacity\right)$
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[H]
	\caption{Constructor \TextGrowthArray}
	\begin{algorithmic}
		\Procedure{$\FuncConstructor$}{$\VarList$}
			\State $\VarList.\FieldHead \gets \FuncNewArray(\VarInitCapacity)$
			\State $\VarList.\FieldTail \gets \FuncNewDynamicArray()$
			\State $\VarList.\FieldSize \gets 0$
			\State $\VarList.\FieldCapacity \gets \VarInitCapacity$
		\EndProcedure
	\end{algorithmic}
\end{algorithm}
